Carbonate chemistry at Ny-Ålesund

### 2022-08-22 awipev-CO2.Rmd
still on air-sea CO2 flux

```{r, echo=FALSE, warning=FALSE, message=FALSE}
d_all <- d_all %>%
  dplyr::mutate(
    pco2_fb_insi_temp = seacarb::pCO2insi(
      pCO2lab = pco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    fco2_fb = seacarb::p2fCO2(
      pCO2 = pco2_fb,
      T = temp_fb,
      Patm = 1,
      P = 0),
    fco2_fb_insi_temp = seacarb::fCO2insi(
      fCO2lab = fco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    pco2_calc_fb_insi_temp = seacarb::pCO2insi(
      pCO2lab = pco2_calc_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    fco2_calc_fb = seacarb::p2fCO2(
      pCO2 = pco2_calc_fb,
      T = temp_fb,
      Patm = 1,
      P = 0),
    fco2_calc_fb_insi_temp = seacarb::fCO2insi(
      fCO2lab = fco2_calc_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m)
) %>% 
  dplyr::select(sort(current_vars()))

@@

# we look at salinity 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(case_when(depth_group == '0-2' | depth_group == '2-4' ~ '0-4', 
                                      depth_group == '>8' ~ '8-12'))) %>% 
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>% 
  dplyr::group_by(depth_group2, montht)
tmp_delta <- tidyr::drop_na(tmp) %>% 
  group_by(depth_group2, montht) %>% 
  dplyr::summarise(sal_med = median(sal_insitu_ctd, na.rm = TRUE)) %>% 
  tidyr::pivot_wider(names_from = depth_group2, values_from = sal_med) %>% 
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>% 
  dplyr::mutate(sal_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median salinity with (median0_4m - median8_12m)")

@@

sal_delta_plt <- tmp_delta %>% 
  ggplot(aes(x=montht, y=sal_delta)) +
  scale_x_discrete("Month") +
  scale_y_continuous("Salinity", limits = c(-1.1, 0.8), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 2, y = 0.65, parse = TRUE, label = "S['surface'] > S['deep']") +
  annotate("text", x = 2, y = -0.9, parse = TRUE, label = "S['surface'] <  S['deep']") +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(sal_delta_plt)
```
### Stratification temperature
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at temperature all depths
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth_group * tmp$montht))
# we look at temperature 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(case_when(depth_group == '0-2' | depth_group == '2-4' ~ '0-4', 
                                      depth_group == '>8' ~ '8-12'))) %>% 
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>% 
  dplyr::group_by(depth_group2, montht)
tmp_delta <- tmp %>% 
  group_by(depth_group2, montht) %>% 
  dplyr::summarise(temp_med = median(temp_insitu_ctd, na.rm = TRUE)) %>% 
  tidyr::pivot_wider(names_from = depth_group2, values_from = temp_med) %>% 
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>% 
  dplyr::mutate(temp_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median temperature with (median0_4m - median8_12m)")
p <- tmp %>% 
  tidyr::drop_na() %>% 
  ggplot(aes(x=depth_group2, y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
temp_delta_plt <- tmp_delta %>% 
  #tidyr::drop_na() %>% 
  ggplot(aes(x = montht, y = temp_delta)) +
  #scale_x_discrete(expand = c(0, 0)) +
  scale_x_discrete("Month") +
  scale_y_continuous("Temperature", limits = c(-2.1, 0.5), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 2, y = 0.3, parse = TRUE, label = "T['surface'] > T['deep']") +
  annotate("text", x = 2, y = -1.8, parse = TRUE, label = "T['surface'] <  T['deep']") +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(temp_delta_plt)
```

@@

pH_delta_plt <- tidyr::drop_na(tmp_delta) %>% 
  ggplot(aes(x=montht, y=pH_tot_sf_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(substitute("pH"["T"] ~ "at 4°C"), limits = c(-0.2, 0.2), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 1.5, y = 0.8, parse = TRUE, label = "pH['surface'] > pH['deep']") +
  annotate("text", x = 1.5, y = -0.15, parse = TRUE, label = "pH['surface'] < pH['deep']") +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
```
### Stratification pCO2
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at pCO2 NOT normalized at 4°C at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                pHint_tot_sf,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(
    at_sf = (intercept_s_at + slope_s_at * sal_insitu_ctd) * 1e-6,
    pco2_sf = seacarb::carb(
      flag = 8,
      var1 = pHint_tot_sf,
      var2 = at_sf,
      S = sal_insitu_ctd,
      T = temp_insitu_ctd,
      Patm = 1,
      P = depth / 10,
      Pt = 0,
      Sit = 0,
      k1k2 = "l",
      kf = "dg",
      ks = "d",
      pHscale = "T",
      b = "u74"
    )$pCO2
  )
tmp <- tmp %>% 
  dplyr::select(montht, pco2_sf, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_sf vs depth", x = "Depth (m)", y= "fCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_sf ~ tmp$depth_group * tmp$montht))

# we look at pCO2 NOT normalized at 4°C 0 to 4 m vs more than 8 m
tmp <- tmp %>%
  dplyr::select(montht, pco2_sf, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
summary(aov(tmp$pco2_sf ~ tmp$depth_group2 * tmp$montht))

tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(pco2_sf_med = median(pco2_sf, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = pco2_sf_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%   
  dplyr::mutate(pco2_sf_med_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 3, caption = "Median seafet pCO2 (median0_4m - median8_12m)")

p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group2), y = pco2_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pCO2 at SeaFET temperature", x = "Depth (m)", y= "pCO2") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

pco2_delta_plt <- tmp_delta %>%
  ggplot(aes(x = montht, y = pco2_sf_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(limits = c(-45,10), expand = c(0, 0)) +
  labs(y = latex2exp::TeX("$pCO_2$ at in situ temperature ($\\mu atm$)")) +
  geom_col(fill = "blue") +
  geom_hline(yintercept = 0) +
  annotate(
    "text",
    x = 2.2,
    y = 6,
    parse = TRUE,
    label = "pCO['2 surface']>pCO['2 deep']"
  ) +
  annotate(
    "text",
    x = 2,
    y = -40,
    parse = TRUE,
    label = "pCO['2 surface']<pCO['2 deep']"
  ) +
  Mytheme(size_labs = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
# we look at pCO2 normalized at 4°C at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                pHint_tot_sf,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(
    at_sf = (intercept_s_at + slope_s_at * sal_insitu_ctd) * 1e-6,
    pco2_sf = seacarb::carb(
      flag = 8,
      var1 = pHint_tot_sf,
      var2 = at_sf,
      S = sal_insitu_ctd,
      T = temp_insitu_ctd,
      Patm = 1,
      P = depth / 10,
      Pt = 0,
      Sit = 0,
      k1k2 = "l",
      kf = "dg",
      ks = "d",
      pHscale = "T",
      b = "u74"
    )$pCO2,
    pco2_sf_4 = seacarb::pCO2insi(pCO2lab = pco2_sf, Tlab = temp_insitu_ctd, SST = 4)
  )
tmp <- tmp %>% 
  dplyr::select(montht, pco2_sf_4, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_sf_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_sf at 4°C vs depth", x = "Depth (m)", y= "pCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_sf_4 ~ tmp$depth_group * tmp$montht))

# we look at pCO2 normalized at 4°C 0 to 4 m vs more than 8 m
tmp <- tmp %>%
  dplyr::select(montht, pco2_sf_4, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
summary(aov(tmp$pco2_sf_4 ~ tmp$depth_group2 * tmp$montht))

tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(pco2_sf_4_med = median(pco2_sf_4, na.rm = FALSE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = pco2_sf_4_med, values_fill = NA) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%   
  dplyr::mutate(pco2_sf_4_med_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 3, caption = "Median seafet pCO2 at 4°C with (median0_4m - median8_12m)")

p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group2), y = pco2_sf_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pCO2 normalized at 4°C", x = "Depth (m)", y= "pCO2") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

pco2_4_delta_plt <- tmp_delta %>%
  ggplot(aes(x = montht, y = pco2_sf_4_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(limits = c(-45,10), expand = c(0, 0)) +
  labs(y = latex2exp::TeX("$pCO_2$ at 4°C ($\\mu atm$)")) +
  geom_col(fill = "blue") +
  geom_hline(yintercept = 0) +
  annotate(
    "text",
    x = 2.2,
    y = 6,
    parse = TRUE,
    label = "pCO['2 surface']>pCO['2 deep']"
  ) +
  annotate(
    "text",
    x = 2.,
    y = -40,
    parse = TRUE,
    label = "pCO['2 surface'] < pCO['2 deep']"
  ) +
  Mytheme(size_labs = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
```

@@

# season 
taka_season <- taka %>% 
  group_by(season) %>% 
  dplyr::mutate(pco2_thermal = 
                  mean(pco2_fb_insi_temp, na.rm = TRUE) * exp(0.0423 * (temp_insitu_11m - mean(temp_insitu_11m, na.rm = TRUE))),
                  pco2_nonthermal = pco2_fb_insi_temp * exp(0.0423 * (mean(temp_insitu_11m, na.rm = TRUE) - temp_insitu_11m))) %>% 
  dplyr::summarise(thermal = max(pco2_thermal, na.rm = TRUE) - min(pco2_thermal, na.rm = TRUE),
                   nonthermal = max(pco2_nonthermal, na.rm = TRUE) - min(pco2_nonthermal, na.rm = TRUE),
                   thermal_nonthermal = thermal / nonthermal) %>% 
  ungroup()
```

@@

```{r air-sea CO2, echo=FALSE, message= FALSE, warning=FALSE}
# calculations as described by de Carlo et al. (2013)
# we consider that mole fraction in air = ppm TO CHECK
# By convention, a negative flux indicates fluxes directed from the atmosphere to the ocean, and a positive flux indicates fluxes directed from the ocean to the atmosphere.

##TO DO: check calculations by Miller et al. (2020); use formulation of Wanninkhof 2014?
## TO DO: Dobashi R. & Ho D. T., 2022. Air-sea gas exchange in a seagrass ecosystem. EGUsphere 
d_all <- d_all %>%
  dplyr::mutate(delta_pco2 = pco2_fb - co2_air, # ppm
                k600 = 0.266 * ws^2, #cm/hr Ho et al. (2006)
                Sc = 2073.1 - (125.62*temp_insitu_11m) + (3.6276*temp_insitu_11m^2) -
                  (0.043219*temp_insitu_ctd^3), #dimensionless
                k = k600 * (Sc/600)^-0.5, #cm/h
                alp = seacarb::K0(S=sal_fb_mix,T=temp_insitu_11m,P=0,Patm=p_air/1013.25), #mol/(kg/atm)
                alp = alp * (seacarb::rho(S = sal_fb_mix, T = temp_insitu_11m, P = 0)/1000), #mol/(l/atm)
                f_air_sea = k * alp * delta_pco2 # mmol/(m2 h)
                )

>>>>

```{r, echo=FALSE, warning=FALSE, message=FALSE}
d_all <- d_all %>%
  dplyr::mutate(
    pco2_fb_insi_temp = seacarb::pCO2insi(
      pCO2lab = pco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    pco2_fb_ctd_temp = seacarb::pCO2insi(
      pCO2lab = pco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_ctd),
    fco2_fb = seacarb::p2fCO2(
      pCO2 = pco2_fb,
      T = temp_fb,
      Patm = 1,
      P = 0),
    fco2_fb_insi_temp = seacarb::fCO2insi(
      fCO2lab = fco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    pco2_calc_fb_insi_temp = seacarb::pCO2insi(
      pCO2lab = pco2_calc_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    fco2_calc_fb = seacarb::p2fCO2(
      pCO2 = pco2_calc_fb,
      T = temp_fb,
      Patm = 1,
      P = 0),
    fco2_calc_fb_insi_temp = seacarb::fCO2insi(
      fCO2lab = fco2_calc_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m)
) %>% 
  dplyr::select(sort(current_vars()))

@@

# we look at salinity 0 to 4 m vs more than 8 m
tmp <- d_all %>%
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
tmp_delta <- tidyr::drop_na(tmp) %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(sal_med = median(sal_insitu_ctd, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = sal_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%
  dplyr::mutate(sal_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median salinity with (median0_4m - median8_12m)")

@@

sal_delta_plt <- tmp_delta %>% 
  ggplot(aes(x=montht, y=sal_delta)) +
  scale_x_discrete("Month") +
  scale_y_continuous("Salinity", limits = c(-1.1, 0.8), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 2, y = 0.65, parse = TRUE, label = "S['surface'] > S['deep']") +
  annotate("text", x = 2, y = -0.9, parse = TRUE, label = "S['surface'] <  S['deep']") +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(sal_delta_plt)
```

### Stratification temperature
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at temperature all depths
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth_group * tmp$montht))
# we look at temperature 0 to 4 m vs more than 8 m
tmp <- d_all %>%
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(temp_med = median(temp_insitu_ctd, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = temp_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%
  dplyr::mutate(temp_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median temperature with (median0_4m - median8_12m)")
p <- tmp %>% 
  tidyr::drop_na() %>% 
  ggplot(aes(x=depth_group2, y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
temp_delta_plt <- tmp_delta %>% 
  #tidyr::drop_na() %>% 
  ggplot(aes(x = montht, y = temp_delta)) +
  #scale_x_discrete(expand = c(0, 0)) +
  scale_x_discrete("Month") +
  scale_y_continuous("Temperature", limits = c(-2.1, 0.5), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 2, y = 0.3, parse = TRUE, label = "T['surface'] > T['deep']") +
  annotate("text", x = 2, y = -1.8, parse = TRUE, label = "T['surface'] <  T['deep']") +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(temp_delta_plt)
```

@@

pH_delta_plt <- tidyr::drop_na(tmp_delta) %>% 
  ggplot(aes(x=montht, y=pH_tot_sf_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(substitute("pH"["T"] ~ "at 4°C"), limits = c(-0.2, 0.2), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 1.5, y = 0.8, parse = TRUE, label = "pH['surface'] > pH['deep']") +
  annotate("text", x = 1.5, y = -0.15, parse = TRUE, label = "pH['surface'] < pH['deep']") +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
```
### Stratification pCO2
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at pCO2 NOT normalized at 4°C at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                pHint_tot_sf,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(
    at_remos = (intercept_s_at + slope_s_at * sal_insitu_ctd) * 1e-6,
    pco2_remos_ctd_temp = seacarb::carb(
      flag = 8,
      var1 = pHint_tot_sf,
      var2 = at_remos,
      S = sal_insitu_ctd,
      T = temp_insitu_ctd,
      Patm = 1,
      P = depth / 10,
      Pt = 0,
      Sit = 0,
      k1k2 = "l",
      kf = "dg",
      ks = "d",
      pHscale = "T",
      b = "u74"
    )$pCO2
  )
# add at_remos and pco2_remos_ctd_temp to d_all
d_all <-
  dplyr::left_join(d_all, dplyr::select(tmp, datetime, at_remos, pco2_remos_ctd_temp), by = "datetime") %>%
  dplyr::select(sort(current_vars()))
tmp <- tmp %>% 
  dplyr::select(montht, pco2_remos_ctd_temp, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_remos_ctd_temp)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_sf vs depth", x = "Depth (m)", y= "fCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_remos_ctd_temp ~ tmp$depth_group * tmp$montht))

# we look at pCO2 NOT normalized at 4°C 0 to 4 m vs more than 8 m
tmp <- tmp %>%
  dplyr::select(montht, pco2_remos_ctd_temp, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
summary(aov(tmp$pco2_remos_ctd_temp ~ tmp$depth_group2 * tmp$montht))

tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(pco2_remos_ctd_temp_med = median(pco2_remos_ctd_temp, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = pco2_remos_ctd_temp_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%
  dplyr::mutate(pco2_remos_ctd_temp_med_delta = median0_4m - median8_12m) %>%
  # this is to bring June back...
  rbind(
    data.frame(
      montht = "Jun",
      median8_12m = NA,
      median0_4m = NA,
      pco2_remos_ctd_temp_med_delta = NA
    )
  ) %>%
  dplyr::mutate(montht = factor(montht, levels = month.abb)) %>%
  dplyr::arrange(montht)
kable(tmp_delta, digits = 3, caption = "Median REMOS pCO2 (median0_4m - median8_12m)")

p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group2), y = pco2_remos_ctd_temp)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pCO2 at SeaFET temperature", x = "Depth (m)", y= "pCO2") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

pco2_delta_plt <- tmp_delta %>%
  ggplot(aes(x = montht, y = pco2_remos_ctd_temp_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(limits = c(-46,10), expand = c(0, 0)) +
  labs(y = latex2exp::TeX("$pCO_2$ at in situ temperature ($\\mu atm$)")) +
  geom_col(fill = "blue") +
  geom_hline(yintercept = 0) +
  annotate(
    "text",
    x = 10,
    y = 6,
    parse = TRUE,
    label = "pCO['2 surface']>pCO['2 deep']"
  ) +
  annotate(
    "text",
    x = 10,
    y = -40,
    parse = TRUE,
    label = "pCO['2 surface']<pCO['2 deep']"
  ) +
  Mytheme(size_labs = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
# we look at pCO2 normalized at 4°C at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                #pHint_tot_sf,
                pco2_remos_ctd_temp,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(pco2_remos_4 = seacarb::pCO2insi(pCO2lab = pco2_remos_ctd_temp, Tlab = temp_insitu_ctd, SST = 4)
  )
# add pco2_remos_4 to d_all
d_all <-
  dplyr::left_join(d_all, dplyr::select(tmp, datetime, pco2_remos_4), by = "datetime") %>%
  dplyr::select(sort(current_vars()))

tmp <- tmp %>% 
  dplyr::select(montht, pco2_remos_4, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_remos_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_remos at 4°C vs depth", x = "Depth (m)", y= "pCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_sf_4 ~ tmp$depth_group * tmp$montht))

# we look at pCO2 normalized at 4°C 0 to 4 m vs more than 8 m
tmp <- tmp %>%
  dplyr::select(montht, pco2_remos_4, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
summary(aov(tmp$pco2_remos_4 ~ tmp$depth_group2 * tmp$montht))

tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(pco2_remos_4_med = median(pco2_remos_4, na.rm = FALSE)) %>%
  tidyr::pivot_wider(names_from = depth_group2,
                     values_from = pco2_remos_4_med,
                     values_fill = NA) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%
  dplyr::mutate(pco2_remos_4_med_delta = median0_4m - median8_12m) %>%
  # this is to bring June back...
  rbind(
    data.frame(
      montht = "Jun",
      median8_12m = NA,
      median0_4m = NA,
      pco2_remos_4_med_delta = NA
    )
  ) %>% 
  dplyr::mutate(montht = factor(montht, levels = month.abb)) %>% 
  dplyr::arrange(montht)
kable(tmp_delta, digits = 3, caption = "Median remos pCO2 at 4°C with (median0_4m - median8_12m)")

p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group2), y = pco2_remos_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pCO2 normalized at 4°C", x = "Depth (m)", y= "pCO2") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

pco2_4_delta_plt <- tmp_delta %>%
  ggplot(aes(x = montht, y = pco2_remos_4_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(limits = c(-46,10), expand = c(0, 0)) +
  labs(y = latex2exp::TeX("$pCO_2$ at 4°C ($\\mu atm$)")) +
  geom_col(fill = "blue") +
  geom_hline(yintercept = 0) +
  annotate(
    "text",
    x = 10,
    y = 6,
    parse = TRUE,
    label = "pCO['2 surface']>pCO['2 deep']"
  ) +
  annotate(
    "text",
    x = 10,
    y = -40,
    parse = TRUE,
    label = "pCO['2 surface'] < pCO['2 deep']"
  ) +
  Mytheme(size_labs = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
```

@@

# season 
taka_season <- taka %>%
  group_by(season) %>%
  dplyr::mutate(
    pco2_thermal =
      mean(pco2_fb_insi_temp, na.rm = TRUE) * exp(0.0423 * (
        temp_insitu_11m - mean(temp_insitu_11m, na.rm = TRUE)
      )),
    pco2_nonthermal = pco2_fb_insi_temp * exp(0.0423 * (
      mean(temp_insitu_11m, na.rm = TRUE) - temp_insitu_11m
    ))
  ) %>%
  dplyr::summarise(
    thermal = max(pco2_thermal, na.rm = TRUE) - min(pco2_thermal, na.rm = TRUE),
    nonthermal = max(pco2_nonthermal, na.rm = TRUE) - min(pco2_nonthermal, na.rm = TRUE),
    thermal_nonthermal = thermal / nonthermal
  ) %>%
  ungroup()
```

@@

## Comparison pCO2 ferrybox vs pCO2 REMOS
pCO2 measured in FerryBox expressed at in situ temp vs pCO2 calculated from REMOS data (S, T, pH, salinity-AT) at 0-2 m only.
